Automated Timed Temporal Verification

### 2022-03-23 tv.tex
m

\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}

@@

    \acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2018}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

@@

\newcommand{\timedEffs}{\emph{TimedEffs}}
\newcommand{\mysharp}{{\mathrel{\texttt{\#}}}}

@@

\usepackage{xcolor}
\definecolor{javablue}{rgb}{0.25,0,1} % for strings
\definecolor{javagreen}{rgb}{0.25,0.5,0.35} % comments
\definecolor{javapurple}{rgb}{0.5,0,0.35} % keywords
\definecolor{javadocblue}{rgb}{0.25,0.35,0.75} % javadoc
\definecolor{mGray}{rgb}{0.4,0.4,0.4}
\definecolor{mPurple}{rgb}{0.58,0,0.82}
\definecolor{huntergreen}{rgb}{0.33, 0.42, 0.18}

\definecolor{darkpastelred}{rgb}{0, 0, 0}
\definecolor{backgroundColour}{rgb}{0.95,0.95,0.92}
\definecolor{brickred}{rgb}{0.8, 0.25, 0.33}
\definecolor{red}{rgb}{0.6,0,0} 
\definecolor{blue}{rgb}{0,0,0.6}
\definecolor{green}{rgb}{0,0.8,0}
\definecolor{cyan}{rgb}{0.0,0.6,0.6}
\definecolor{cloudwhite}{rgb}{0.9412, 0.9608, 0.8471} 
\definecolor{davysgrey}{rgb}{0.33, 0.33, 0.33}
\definecolor{deepfuchsia}{rgb}{0.76, 0.33, 0.76}
\definecolor{deeplilac}{rgb}{0.6, 0.33, 0.73}
\definecolor{deepskyblue}{rgb}{0.0, 0.75, 1.0}
\definecolor{indianred}{rgb}{0.8, 0.36, 0.36}
\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
\definecolor{darkred}{rgb}{0.55, 0.0, 0.0}
\usepackage{wrapfig}
\newcommand{\effect}{\textcolor{black}{\ensuremath{\mathrm{\Phi}}}}
\definecolor{color_pure}{RGB}{3,35,14}
\newcommand\pure[1]{ \textcolor{black}{#1}}
\usepackage{epigraph}
\newcommand{\anyevent}[1]{{\textcolor{darkred}
{\underline{\textbf{\footnotesize #1}}}}}
\newcommand{\seq}{\cdot}
\newcommand{\choice}{\vee}
\newcommand{\fullname}{Integrated Dependent Effects}
\newcommand{\name}{integrated effects}
\newcommand{\code}[1]{{\tt{\ensuremath{\m{#1}}}}}
\newcommand{\codeme}[1]{{\tt{\ensuremath{#1}}}}
\newcommand{\esn}[2]{\ensuremath{#1^{#2}}}
\newcommand{\empt}{\textcolor{black}{\ensuremath{\epsilon}}}
\newcommand{\bott}{\textcolor{black}{\ensuremath{\bot}}}
\newcommand{\CONTAIN}{\sqsubseteq}
\newcommand{\underscore}{\textcolor{vividauburn}{\code{\_}}}
\newcommand{\m}{\mathit}  
\newcommand{\toolName}{Tvide}
\def\defeq{\ensuremath{\,\triangleq}}
\definecolor{regalia}{rgb}{0.32, 0.18, 0.5}
\newcommand{\inclusion}{\code{\mathcal{I}}}
\definecolor{mypink}{RGB}{219, 48, 122}
\definecolor{darklavender}{rgb}{0.45, 0.31, 0.59}
\definecolor{deepcerise}{rgb}{0.85, 0.2, 0.53}
\newcommand\figref[1]{Fig. \textcolor{black}{\ref{#1}}.}
\newcommand\tabref[1]{Table \textcolor{black}{\ref{#1}}.}
\newcommand\secref[1]{Sec. \textcolor{black}{\ref{#1}}}

@@

\lstdefinelanguage{JavaScript}{
  keywords={typeof, const, new, true, false, catch, function, return, null, catch, switch, var,async, await, if, setTimeout, this, then, while, post, else, case, break, return},
  keywordstyle=\color{blue}\bfseries,
  ndkeywords={class, export, boolean, throw, implements, import, this},
  ndkeywordstyle=\color{darkgray}\bfseries,
  identifierstyle=\color{black},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{darklavender}\ttfamily,
  stringstyle=\color{red}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract} 
To make reactive programming more concise and flexible, it is promising to deploy a mixed concurrency paradigm \cite{berry2020hiphop} that
integrates Esterel's synchrony and preemption \cite{berry1999constructive}  
with JavaScript's asynchrony \cite{madsen2017model}. 
%HipHop.js \cite{berry2020hiphop} is the first web language that 
Existing temporal verification techniques have not been designed to handle such a blending of two concurrency
models.
We propose a novel solution via a 
compositional 
Hoare-style forward verifier and a term rewriting system (TRS) on \emph{Timed Synchronous Effects} (TSE).

Firstly, we introduce TSE, a new effects logic, that extends \emph{Kleene Algebra} with value-dependent constraints, providing real-time bounds for logical-time synchronous traces \cite{von2017real}. 
Secondly, we establish an axiomatic semantics  for a core language \code{\lambda_{HH}}, generalising the mixed paradigm. 
% HipHop.js,  based on our effects logic. 
%Secondly, we formalise a set of construction rules to infer the program effects based on a . using a set of inductive rules
Thirdly, we present a purely algebraic TRS, to efficiently check language inclusions between expressive timed effects.  
To demonstrate the feasibility of our proposals, we prototype the  verification system; prove its correctness; investigate how it can help to debug errors related to both synchronous and asynchronous programs.

@@

\ccsdesc[500]{Theory of computation~Logic}
\ccsdesc[300]{Theory of computation~Semantics and reasoning}
\ccsdesc[100]{Models of computation~Concurrency}

@@

As it recently proposed,  reactive languages such as Hiphop.js\footnote{Hiphop.js is a JavaScript extension of Esterel \cite{berry1992esterel}
(or vice versa)
for reactive web applications: \url{https://www-sop.inria.fr/members/Colin.Vidal/hiphop/}.}   \cite{vidal2018hiphop,berry2020hiphop} are designed to be
based on a smooth integration of (i) Asynchronous concurrent programs, which perform interactions between components or with the environment with uncontrollable timing, such as network-based communication; (ii) Synchronous reactive programs, which react to external events in a conceptually instantaneous and deterministic way; and (iii) Preemption, the explicit cancellation and resumption of an ongoing orchestration subactivity.

%Recent research has

%One of the major difficulties of reactive web programming is \emph{callback hell} \cite{kambona2013evaluation}, which is the problem of appropriately handling the asynchronous events appearing in program executions. To avoid using callbacks, efforts have been made: 
%%\cite{madsen2017model}
%(i) JavaScript has provided  \emph{promises} and \emph{async/await} primitives, making it possible to chain asynchronous actions in a specific sequential order; (ii) High-level abstractions such as \emph{Functional Reactive Programming} has been carried up to the web applications \cite{czaplicki2013asynchronous}, which adopts the data flow declarative programming style: when a variable is modified, any expression that references it is implicitly reevaluated. 
%
%To go further with this problem and achieve a better web orchestration,  the language Hiphop.js \cite{vidal2018hiphop,berry2020hiphop} is designed to be a JavaScript extension of Esterel \cite{berry1992esterel} (or vice versa) based on a smooth integration of (i) Asynchronous concurrent programs, which perform interactions between components or with the environment with uncontrollable timing, such as network-based communication; (ii) Synchronous reactive programs, which react to external events in a conceptually instantaneous and deterministic way; and (iii) Preemption, the explicit cancellation and resumption of an ongoing orchestration subactivity.

\textit{"Such a combination makes reactive programming more powerful and flexible than plain JavaScript because it makes the temporal geometry of complex executions explicit instead of hidden in implicit relations between state variables." \cite{berry2020hiphop}}

Given such a multi-paradigm (mixed Sync-Async) concurrency model, the verification of its temporal behaviour becomes engaging and challenging. 
%which, however, has not been intensely exploited. 
Existing techniques are based on a transformation to convert the asynchronous chunk into semantically equivalent synchronization\footnote{Usually the transformation is incomplete, i.e., not all the asynchronous programs can be translated into a semantically equivalent synchronization.}; then reason about the behaviours based on the synchronous semantics \cite{damian2019communication,gleissenthall2019pretend,tarawneh2018formal}. This approach (i) suffers from the limited expressiveness, restricted by finite-state automata (FSA); (ii) lacks the modularity to reason about programs compositionally; (iii) loses time awareness due to the conversion to synchronous models; and (iv) has the \emph{state explosion problem}  when proving language inclusions between FSA.
 %finite-state automata
 % PSPACE-complete complexity

@@

\begin{enumerate}
\item \textbf{The Timed Synchronous Effects:} We define the syntax ({\secref{subsec:Specification_language}}) and semantics ({\secref{subsec:Specification_Semantics}}) of timed synchronous effects, to be the specification language, which captures the target programs' behaviours and non-trivial temporal properties.

@@

\noindent\textbf{\textit{Organization.}} 
\secref{sec:Preliminaries} introduces the language features of synchronous Esterel programs, JavaScript promises, and the web reactive language HipHop.js. 
\secref{sec:Overview} gives motivation examples to highlight the key methodologies and contributions. 
\secref{sec:LanguageSpecifications} formally presents the core   language \code{\lambda_{HH}}, and the syntax and semantics of timed synchronous effects.
\secref{sec:Verification} presents the forward verification rules.
\secref{sec:Entailment_Prover} explains the TRS for effects inclusion checking, and displays the essential auxiliary functions. 
\secref{sec:Evaluation} demonstrates the implementation and cases studies. % via a number of . 
We discuss related works in \secref{sec:Related_work} and conclude in  \secref{sec:conclusion}. Omitted proofs can be found in the Appendix.

\section{Background: Esterel, Async-Await, and Hiphop.js }
\label{sec:Preliminaries}

It has been an active research topic to build flexible programming paradigms for reactive systems in different domains. This section identifies: i) \emph{Synchronous programming}, represented by Esterel  \cite{berry1992esterel}, ii) \emph{Asynchronous programming}, represented by JavaScript promises \cite{madsen2017model}, and iii) A \emph{mixed Sync-Async} paradigm, recently proposed by HipHop.js  \cite{berry2020hiphop}. Meanwhile, we discuss the real-world programming challenges of each paradigm; and show how our proposal addresses them in the cases studies, \secref{subsec:Case_Studies}.

 %language abstractions

\subsection{A Sense of Esterel: Synchronous and Preemptive}

%Synchronous programming (also synchronous reactive programming or SRP) is a computer programming paradigm supported by synchronous programming languages.

The principle of synchronous programming is to design a high-level abstraction where the timing characteristics of the electronic transistors are neglected \cite{esterel_wiki}. Such an abstraction makes reasoning about time a lot simpler, thanks to the notion of \emph{logical ticks}: a synchronous program reacts to its environment in a sequence of ticks, and computations within a tick are assumed to be instantaneous. 
As one of the first few well-known synchronous languages, Esterel's high-level imperative style allows the simple expression of parallelism and preemption, making it natural for programmers to specify and reason about control-dominated model designs. Esterel has found success in many safety-critical applications such as nuclear power plant control software. The success with real-time and embedded systems in domains that need strong guarantees can be partially attributed to its precise semantics and computational model  \cite{berry1992esterel,berry1999constructive, song2021synchronous}.

Esterel treats computation as a series of deterministic reactions to external signals. All parts of a reaction complete in a single, discrete-time step called an \emph{instance}. Besides, instances exhibit deterministic concurrency; each reaction may contain concurrent threads without execution order affecting the computation result. Primitive constructs execute in zero time except for the \code{yield} (pause) statement. Hence, time flows as a sequence of logical instances separated by explicit pauses. In each instance, several elementary instantaneous computations take place simultaneously.

\begin{lstlisting}
fork {emit A; yield; emit B; emit C} par {emit E; yield; emit F; yield; emit G}
\end{lstlisting}

The synchronous parallelism in Esterel is constructed by the \code{\m{fork} \{...\} \m{par}  \{...\}  } statement. 
It remains active as long as one of its branches remains active, 
% unless some branch exits a trap, 
and it terminates when both branches are terminated. The branches can terminate in different instances, and wait for the last one to terminate.
As the above example shows, the first branch generates effects $\{\textbf{A}\} \cdot \{\textbf{B}, \textbf{C}\}$ while the second branch generates effects $\{\textbf{E}\} \cdot \{\textbf{F}\} \cdot \{\textbf{G}\}$; then the final effects should be $\{\textbf{A},\textbf{E}\}\cdot \{\textbf{B}, \textbf{C}, \textbf{F}\} \cdot \{\textbf{G}\}$.

To maintain determinism and synchrony, evaluation in one thread of execution may affect code arbitrarily far away in the program. In other words, there is a strong relationship between signal status and control propagation: a signal status determines which branch of a \code{present} test is executed, which in turn determines which \code{emit} statements are executed (See \secref{subsec:Targetlanguage} for the language syntax). 
%For instance, selecting a particular branch of execution in one thread may immediately unblock a different thread of execution. 
%The selection of the other branch may render the entire program invalid\citeref{florence2019calculus}. 
The first semantic challenge of programming Esterel is the \emph{Logical Correctness} issue, caused by these non-local executions, which is simply the requirement that there exists precisely \textbf{one} status for each signal. For example, consider the program below:

\begin{lstlisting}
signal S1 in present S1 then nothing else emit S1 end present end signal
\end{lstlisting}

If the local signal {\textbf{S1}} were \emph{present}, the program would take the first branch of the condition, and the program would terminate without having emitted {\textbf{S1}} (\code{nothing} leaves \code{\textbf{S1}} with   \emph{absent}). If \code{\textbf{S1}} were absent, the program would choose the second branch and emit the signal. Both executions lead to a contradiction. Therefore there are no valid assignments of signals in this program. This program is logically incorrect.

\begin{lstlisting}
signal S1 in present S1 then emit S1 else nothing end present end signal
\end{lstlisting}

Consider the revised program above. If the local signal {\textbf{S1}} were present, the conditional would take the first branch, and  {\textbf{S1}} would be emitted, justifying
the choice of signal value. 
If {\textbf{S1}} were absent, the signal would not be emitted, and the choice of absence is also justified. Thus there are two possible assignments to the signals in this program, which is also logically incorrect.

\begin{wrapfigure}{R}{0.41\columnwidth}
    \vspace{-1mm}
\centering
\begin{lstlisting}
module a_loop: output A,B,C;
  emit A;
  loop 
  	emit B; yield; emit C
  end loop  
end module
\end{lstlisting}  
     \vspace{-1mm}
 \caption{A Loop Example in Esterel \cite{song2021synchronous}.} 
  \label{fig:loop1}
       \vspace{-1mm}
\end{wrapfigure}

Esterel's instantaneous nature requires a special distinction when it comes to loop statements, which increases the difficulty of the effects invariants inference.  
As shown in \figref{fig:loop1}, the program firstly emits signal {\textbf{A}}, then enters into a loop which emits signal \code{\textbf{B}} followed by a \code{\m{yield}} followed by emitting signal {\textbf{C}} at the end. 
The effects of it is $ \{\textbf{A}, \textbf{B}\} \cdot \{ \textbf{B}, \textbf{C}\} \cdot \{ \textbf{B}, \textbf{C}\} \cdot \{ \textbf{B}, \textbf{C}\} ...$,
which says that in the first time instance, signals {\textbf{A}} and {\textbf{B}} will be present, 
as there is no explicit yield between $emit\ \textbf{A}$ and $emit\ \textbf{B}$; 
then for the following instances (in an infinite trace), signals {\textbf{B}} and {\textbf{C}} are present all the time, because after executing $emit\ \textbf{C}$, it immediately executes from the beginning of the loop, which is $emit\ \textbf{B}$.

 %one deals with concurrent processes that coordinate with each other.

Coordination in concurrent systems can result from information exchange, using messages circulating on channels with possible implied synchronization. In Esterel, it can also result from \emph{process preemption}  \cite{berry1993preemption}, which is a more implicit control mechanism that consists in denying the right to work to a process, either permanently (abortion, cf. \figref{fig:overview_eg1}) or temporarily (e.g. suspension). Preemption is particularly important in control-dominated reactive and real-time programming, where most of the works consist of handling interrupts and controlling computation.

%
%\begin{wrapfigure}{L}{0.4\columnwidth}
%    \vspace{-1mm}
%\centering
%\begin{lstlisting}
%suspend
%	loop emit A; yield; yield;
%	end loop
%when B end suspend
%\end{lstlisting}  
%     \vspace{-1mm}
% \caption{A Suspend Example in Esterel.} 
%  \label{fig:Suspend}
%       \vspace{5mm}
%\includegraphics[width=0.39\columnwidth]{suspend.png}
%        \vspace{-1mm}
%\caption{\label{fig:Suspend}Graphical Illustration of \figref{fig:Suspend}}
%        \vspace{0mm}
%\end{wrapfigure}
%
%
%\figref{fig:Suspend} presents a suspend statement, one of the preemptive operations in Esterel. When the preemption condition satisfied, it pauses the execution of a group of statements, and resume it from the next time instance. 
%As \figref{fig:Suspend_pbg} illustrates, without the preemption, the effects is a repeated pattern: \code{\{\textbf{A}\}\cdot\{\}\cdot\{\textbf{A}\}\cdot\{\} \cdot ...} . Due to the presence of signal \code{\textbf{B}}, it delays emission of \code{\textbf{A}} by one cycle (at {\textcircled{1}} and {\textcircled{2}}), and resumes from the next cycle.

While the flexibility they provide us, preemption primitives often with loose or complex semantics, making abstract reasoning difficult. 
Most existing languages offer a small set of preemption primitives, often insufficient to program reactive systems concisely.

\subsection{Asynchrony from JavaScript Promises: Async–Await}

\noindent \textit{"Who can wait quietly while the mud settles? Who can remain still until the moment of action?"}\\
\textcolor{white}{.}\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad \textit{ -- Laozi, Tao Te Ching}

A number of mainstream languages, such as C\code{\#}, JavaScript, Rust, and Swift, have recently added support for async–await and the accompanying promises abstraction\footnote{JavaScript's asynchrony arises in situations such as web-based user-interfaces, communicating with servers through HTTP requests, and non-blocking I/O.}, also known as \emph{futures} or \emph{tasks} \cite{bierman2012pause}. As an example, consider the JavaScript program in \figref{fig:json}, 
it uses the {\ttfamily{fs}} module (line 1) to load the file into a variable (line 6) using async/await syntax.

\begin{figure}[h]
    \vspace{-2mm}
\centering
\begin{lstlisting}
const fs = require('fs').promises;

async function read (filePath) {
	const task = fs.readFile(filePath);
    ... // do things that do not depend on the result of the loading file
    const data = await task; // block execution until the file is loaded
	... // logging or data processing of the Json file 
}
\end{lstlisting}  
     \vspace{-3mm}
 \caption{Using Async-Await in JavaScript.} 
  \label{fig:json}
       \vspace{-2mm}
\end{figure}

The function {\ttfamily{read}} accepts one argument, a string called {\ttfamily{filePath}}. As it is declared {\textbf{\color{blue}\ttfamily{async}}}, reading a file (line 4) does not block computations that do not depend on the result (line 5). The programmer {\textbf{\color{blue}\ttfamily{await}}}s the task when they need the result to be ready. Reading a file may raise exceptions (e.g., the file is non-existent), then the point for such an exception to emerge is where the tasks are awaited (lines 6). While await sends a signal to a task scheduler on the runtime stack, the exceptions appear to propagate in the opposite direction, from the runtime to the await sites.

Unfortunately, existing languages that support async–await do not enforce at compile time that exceptions raised by asynchronous computations are handled. The lack of this static assurance makes asynchronous programming error-prone. For example, the JavaScript compiler accepts the program above without requiring that an exception handler to be provided, which leads to program crashes if the asynchronous computation results in an exception. 
%The worse situation is: an exception raised asynchronously is possibly silently swallowed if not otherwise caught. 
Such unhandled exceptions have been identified as a common vulnerability in JavaScript programs \cite{madsen2017model,alimadadi2018finding}.
Furthermore, \cite{madsen2017model,alimadadi2018finding} display a set of other \emph{anti-patterns} including: 
attempting to settle a promise multiple times; unsettled promises; unreachable reactions; 
%implicit returns in reactions; 
and unnecessary promises, which are mostly caused by using promises without sufficient static checking.

   \vspace{2mm}
\subsection{HipHop.js -- A mixture of Esterel and JavaScript}
\label{subsection:HipHop}

\begin{figure}[h]
\noindent
   \vspace{-2mm}
\begin{varwidth}{.5\textwidth}
\begin{lstlisting}
function enableLoginButton(){
  return (Rname.length >= 2 
	   && Rpasswd.length >= 2);}
function nameKeypress(value){
  Rname = value;
  RenableLogin=enableLoginButton();}
function passwdKeypress(value){ 
  Rpasswd = value;
  RenableLogin=enableLoginButton();}
\end{lstlisting}
\end{varwidth}%
\begin{varwidth}{.5\textwidth}
\begin{lstlisting}
hiphop module Identity(
			in name, in passwd, 
			out enableLogin){
  do{
	  emit enableLogin(
		  		name.length >= 2 
		  	 && passwd.length >= 2); 
  } every(name || passwd)
}
\end{lstlisting}  
\end{varwidth}
      \vspace{0mm}
\caption{A comparison between JavaScript (left) and HipHop.js (right) for a same login button implementation \cite{berry2020hiphop}. (On the right, {\ttfamily{name}}, {\ttfamily{passwd}} and {\ttfamily{enableLogin}} are reactive input/output signals.)}
  \label{fig:login_compare}  
      \vspace{0mm}
\end{figure}

HipHop.js is a reactive web language that adds synchronous concurrency and preemption to JavaScript, which is compiled into plain  JavaScript and executes on runtime environments \cite{berry2020hiphop}. 
To show the advantages of such a mixture, \figref{fig:login_compare} presents a comparison between JavaScript and HipHop.js to achieve the same login button. Here, {\ttfamily{Rname}}, {\ttfamily{Rpasswd}}, {\ttfamily{RenableLogin}} are global variables to model the application's states. Dis/En-abling login is done by setting {\ttfamily{RenableLogin}}. However, while more and more features get added to the specification, state variable interactions can lead to a large number of implicit and  invisible global control states.

Whereas, HipHop.js simplifies and modularizes designs, and synchronous signaling makes it possible to instantly communicate between concurrent statements to exchange data and coordination signals. 
Second, powerful event-driven reactive preemption borrowed from Esterel finely controls the lifetime of the arbitrarily complex program statements they apply, instantly killing them when their control events occur. 
More examples are discussed in the following sections.

%Third, an extension of async that synchronously signals termination completion to concurrent statements and automatically performs automatic resource cleaning when preempted.

@@

\begin{figure}[h]
      \vspace{0mm}
\begin{lstlisting}[columns=fullflexible]
hiphop module authenticate(var d, var name, var passwd, in Tick, out Connecting, out Connected)
/*@ requires d>3000 : {}^*.{Login} @*/
/*@ ensures	(3000<t/\t<d : {Connecting}#t.{Connected}) \/ (t=d :  {Connecting}#t) @*/
{
   abort count(d, Tick) { // Abort the execution at d milliseconds
      async Connected {
         emit Connecting; 
         // Execute authenticateSvc after a 3000 milliseconds' delay
         setTimeout (authenticateSvc(name, passwd).post().then( v => this.notify(v)), 3000);}}}
\end{lstlisting}  
      \vspace{0mm}
      \caption{Dependent Values for Time Bounds (the program is from  \cite{berry2020hiphop}).}\label{fig:overview_eg1}
         \vspace{0mm}
\end{figure}

@@

{

\begin{figure}[!ht]
\renewcommand{\arraystretch}{1.2}
\centering
  $
  \begin{array}{lrcl}

    \m{(Program)} &  \code{\mathcal{P} }&\code{  ::= }&
    \code{%{datat^*}  \  
   {meth^*}} 
\qquad   \qquad
    \m{(Basic\ Types)} \quad  \code{\tau} \code{\ \ ::= \ \ }
     \code{int\ 
    % |\ bool\ 
     |\ void}
    \\

@@

%     \m{(Types)} &  \code{t}&\code{  ::= }&
% \code{d\ |\ \tau} 
% \\    
  \m{(Method\ Def.)} &  \code{meth}&\code{  ::= }&
  \code{\tau \ mn  \ {( \tau\ x)^*}\ \{\textbf{requires}\ \effect_{pre}\ \textbf{ensures}\  \effect_{post}\} \ \{ e \}}
  \\
  \code{(Values)} &  {v}&\code{  ::= }&
  ()\ |\ c \ | \  x

 \\
  \m{(Expressions)} &  \code{e}&\code{  ::= }&
    \code{ v\ 
    |\ mn({x^*}) \ 
    |\ \textbf{let}\ x{=}e_1\ \textbf{in} \ e_2
    |\ \textbf{if}\ v\ \textbf{then}\ e_1\ \textbf{else}\ e_2 
 }\\ %|\ \textbf{assert}\ \effect  }  \\
 &   & & \code{
  |\ \textbf{emit} [\anyevent{a}]\ 
  | \  \textbf{timeout}\ e \ d \ | \ \textbf{abort} \ e \ d 
  }  \\~\\

>>>>

%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
\documentclass[acmsmall,10pt,review]{acmart}

@@

    \acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2018}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

@@

@@

\newcommand{\timedEffects}{\emph{TimEffs}}
\usepackage{xcolor}
\definecolor{javablue}{rgb}{0.25,0,1} % for strings
\definecolor{javagreen}{rgb}{0.25,0.5,0.35} % comments
\definecolor{javapurple}{rgb}{0.5,0,0.35} % keywords
\definecolor{javadocblue}{rgb}{0.25,0.35,0.75} % javadoc
\definecolor{mGray}{rgb}{0.4,0.4,0.4}
\definecolor{mPurple}{rgb}{0.58,0,0.82}
\definecolor{huntergreen}{rgb}{0.33, 0.42, 0.18}

\definecolor{darkpastelred}{rgb}{0, 0, 0}
\definecolor{backgroundColour}{rgb}{0.95,0.95,0.92}
\definecolor{brickred}{rgb}{0.8, 0.25, 0.33}
\definecolor{red}{rgb}{0.6,0,0} 
\definecolor{blue}{rgb}{0,0,0.6}
\definecolor{green}{rgb}{0,0.8,0}
\definecolor{cyan}{rgb}{0.0,0.6,0.6}
\definecolor{cloudwhite}{rgb}{0.9412, 0.9608, 0.8471} 
\definecolor{davysgrey}{rgb}{0.33, 0.33, 0.33}
\definecolor{deepfuchsia}{rgb}{0.76, 0.33, 0.76}
\definecolor{deeplilac}{rgb}{0.6, 0.33, 0.73}
\definecolor{deepskyblue}{rgb}{0.0, 0.75, 1.0}
\definecolor{indianred}{rgb}{0.8, 0.36, 0.36}
\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
\definecolor{darkred}{rgb}{0.55, 0.0, 0.0}
\usepackage{wrapfig}
\newcommand{\effect}{\textcolor{black}{\ensuremath{\mathrm{\Phi}}}}
\definecolor{color_pure}{RGB}{3,35,14}
\newcommand\pure[1]{ \textcolor{black}{#1}}
\usepackage{epigraph}
\newcommand{\anyevent}[1]{{\textcolor{darkred}
{\underline{\textbf{\footnotesize #1}}}}}
\newcommand{\seq}{\cdot}
\newcommand{\choice}{\vee}
\newcommand{\fullname}{Integrated Dependent Effects}
\newcommand{\name}{integrated effects}
\newcommand{\code}[1]{{\tt{\ensuremath{\m{#1}}}}}
\newcommand{\codeme}[1]{{\tt{\ensuremath{#1}}}}
\newcommand{\esn}[2]{\ensuremath{#1^{#2}}}
\newcommand{\empt}{\textcolor{black}{\ensuremath{\epsilon}}}
\newcommand{\bott}{\textcolor{black}{\ensuremath{\bot}}}
\newcommand{\CONTAIN}{\sqsubseteq}
\newcommand{\underscore}{\textcolor{vividauburn}{\code{\_}}}
\newcommand{\m}{\mathit}  
\newcommand{\toolName}{Tvide}
\def\defeq{\ensuremath{\,\triangleq}}
\definecolor{regalia}{rgb}{0.32, 0.18, 0.5}
\newcommand{\inclusion}{\code{\mathcal{I}}}
\definecolor{mypink}{RGB}{219, 48, 122}
\definecolor{darklavender}{rgb}{0.45, 0.31, 0.59}
\definecolor{deepcerise}{rgb}{0.85, 0.2, 0.53}
\newcommand\figref[1]{Fig. \textcolor{black}{\ref{#1}}.}
\newcommand\tabref[1]{Table \textcolor{black}{\ref{#1}}.}
\newcommand\secref[1]{Sec. \textcolor{black}{\ref{#1}}}
\newcommand{\timedL}{\code{\lambda^{t}}}

@@

\lstdefinelanguage{JavaScript}{
  keywords={typeof, void, const, new, int, true, false, catch, function, return, null, catch, switch, var,async, await, if, setTimeout, this, then, while, post, else, case, break, return},
  keywordstyle=\color{blue}\bfseries,
  ndkeywords={class, export, boolean,event, timeout, deadline, delay,  throw, implements, import, this},
  ndkeywordstyle=\color{darkgray}\bfseries,
  identifierstyle=\color{black},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{darklavender}\ttfamily,
  stringstyle=\color{red}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract} 
In this work, we study temporal verification of 
compositional real-time systems, which are modeled 
using mutable variables and timed processes.  
Instead of explicitly manipulating clock variables, 
a number of compositional timed behavioral patterns, 
such as \emph{delay}, \emph{timeout}, \emph{deadline},  
are introduced to capture quantitative timing constraints.
The idea is to dynamically create clocks, 
and solve constraints on the clocks.

We propose a novel solution for timed verification via a 
compositional Hoare-style forward verifier and a 
term rewriting system (TRS) on \emph{Timed Effects} (\timedEffects).  
More specifically, we formally define 
a core language \timedL, generalizing the timed systems.
Secondly, we introduce \timedEffects, a new effects logic, 
that extends 
\emph{Kleene Algebra} with value-dependent real-time constraints. 
Thirdly,  we establish an  axiomatic semantics for  
\timedL\  to infer temporal behaviors  of  given 
programs, expressed in \timedEffects. 
Lastly, we present a purely algebraic TRS, to 
efficiently prove  language inclusions between 
 \timedEffects.
To demonstrate the feasibility of our proposals, 
we prototype the verification system; prove its 
correctness; report case studies and the experimental results. 
%we prototype the  verification system in OCaml; then prove soundness of both the effects inference process and the term rewriting process. % in Coq.

@@

\ccsdesc[500]{Theory of computation~Logic}
\ccsdesc[300]{Theory of computation~Semantics and reasoning}
\ccsdesc[300]{Theory of computation~Finite Model Theory}

@@

Existing techniques are based on a transformation to convert the asynchronous chunk into semantically equivalent synchronization\footnote{Usually the transformation is incomplete, i.e., not all the asynchronous programs can be translated into a semantically equivalent synchronization.}; then reason about the behaviours based on the synchronous semantics \cite{damian2019communication,gleissenthall2019pretend,tarawneh2018formal}. This approach (i) suffers from the limited expressiveness, restricted by finite-state automata (FSA); (ii) lacks the modularity to reason about programs compositionally; (iii) loses time awareness due to the conversion to synchronous models; and (iv) has the \emph{state explosion problem}  when proving language inclusions between FSA.
 %finite-state automata
 % PSPACE-complete complexity

@@

\begin{enumerate}
\item \textbf{The Timed Effects:} We define the syntax ({\secref{subsec:Specification_language}}) and semantics ({\secref{subsec:Specification_Semantics}}) of timed synchronous effects, to be the specification language, which captures the target programs' behaviours and non-trivial temporal properties.

@@

@@

\begin{figure}[h]
      \vspace{0mm}
\begin{lstlisting}[columns=fullflexible]
  void addSugar(int n)
  // requires n>0 /\ Cup 
  // ensures  t>=n /\ Sugar # t
  {
      if n == 0 then event[Sugar] 
      else {
          timeout (oneSugar(), 1);
          addSugar (n-1);
      }
  }

  void makeCoffee (int n)
  // requires n>0 /\ emp 
  // ensures  5>t>=n/\ t1<3 /\ Cup.(Sugar # t).(Coffee # t1).Done
  {
      event[Cup];
      deadline (getSugar(n), 5);
      deadline (event[Coffee], 3);
      event[Done];
  }

  void main () 
  // requires true /\ emp 
  // ensures  t < 8 /\ (_^*.Done) # t
  {
      makeCoffee(3);
  }
\end{lstlisting}  
      \vspace{0mm}
      \caption{Dependent Values for Time Bounds (the program is from  \cite{berry2020hiphop}).}\label{fig:overview_eg1}
         \vspace{0mm}
\end{figure}

@@

{
 % \vspace{-2mm}
\begin{figure}[!ht]
\renewcommand{\arraystretch}{1.1}
\centering%\small
  $
  \begin{array}{lrcl}

    \m{(Program)} &  \code{\mathcal{P} }&\code{  ::= }&
    \code{%{datat^*}  \  
   {meth^*}} 
\qquad   \qquad
    \m{(Basic\ Types)} \qquad  \code{\tau} \code{  ::= }
     \code{int\ |\ bool\ |\ void}
    \\

@@

%     \m{(Types)} &  \code{t}&\code{  ::= }&
% \code{d\ |\ \tau} 
% \\    
  \m{(Method\ Def.)} &  \code{meth}&\code{  ::= }&
  \code{\tau \ mn  \ {( \tau\ x)^*}\ \{\textbf{requires}\ \effect_{pre}\ \textbf{ensures}\  \effect_{post}\} \ \{ e \}}
  \\
  \m{(Expressions)} &  \code{e}&\code{  ::= }&
    \code{ ()\ |\ k^\tau\ |\ x\ |\ 
    %v.f\ |\ 
    %\textbf{new}\ d({v^*}) \ |\ 
    \tau\ x;\ e \ |\ mn({x^*}) \ }
  \code{|\ x{:=}e\ |\ 
 % v.f:=e \ | 
 \ e_1;e_2\ 
  |\ {if}\ v\ {then}\ e_1\ {else}\ e_2 
 }  \\
 &   & & \code{
% |\ e_1}\ op\ \code{ e_2 \ 
| \ \textbf{event} [\anyevent{A}]\
  %|\ \textbf{assert}\ \effect \ 
  | \ \textbf{timeout} \ e \ d \  
  | \ \textbf{deadline} \ e \ d \  
  | \ \textbf{delay}\ d 
  }  \\
  ~\\
